---
title: "Exercise, Penalized Regression"
author: "Ian"
date: "2023/02/27"
output:
  word_document: default
  html_document: default
---

```{r}
library(MASS)
library(tidyverse)
library(caret)
library(glmnet)
library(caTools)
```

Question 1

```{r}
data("Boston", package = "MASS")
str(Boston) # the response variable is 'medv' and the other 13 are used to fit the model
set.seed(123)
training.samples <- Boston$medv %>% createDataPartition(p = 0.75, list = FALSE)
train.data  <- Boston[training.samples, ]
test.data <- Boston[-training.samples, ]
str(train.data) #381 obs
str(test.data) #125 obs
```

Question 2

```{r}
x <- model.matrix(medv~., train.data)[,-1]
y <- train.data$medv
cv <- cv.glmnet(x, y, alpha = 0) # Does k-fold cross-validation for glmnet, produces a plot, and returns a value for lambda (and gamma if relax=TRUE)
cv$lambda.min
```

Best lambda for ridge regression is 0.6490823

```{r}
model <- glmnet(x, y, alpha = 0, lambda = cv$lambda.min) # alpha=0: ridge
coef(model)
```


```{r}
x.test <- model.matrix(medv ~., test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
data.frame(
  RMSE = RMSE(predictions, test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```
RMSE is 6.635525 and coefficient of determination is 0.6626213.

Question 3
```{r}
cv <- cv.glmnet(x, y, alpha = 1)
cv$lambda.min
```

Best lambda for LASSO is 0.02943509

```{r}
model <- glmnet(x, y, alpha = 1, lambda = cv$lambda.min) # alpha=1: lasso
coef(model)
```

```{r}
x.test <- model.matrix(medv ~., test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
```

```{r}
data.frame(
  RMSE = RMSE(predictions, test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```
RMSE is 6.472275 and coefficient of determination is 0.6749717.

Question 4
```{r}
model <- train(
  medv ~., data = train.data, method = "glmnet",
  trControl = trainControl("cv", number = 10),
  tuneLength = 10
)
model$bestTune
```

```{r}
coef(model$finalModel, model$bestTune$lambda)
```

```{r}
x.test <- model.matrix(medv ~., test.data)[,-1]
predictions <- model %>% predict(x.test)
data.frame(
  RMSE = RMSE(predictions, test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```
RMSE is 6.433159 and coefficient of determination is 0.6781032

Different way spliting data and randomForest function
**Use the following function will get a different result. Both are correct.**

Question 1

```{r}
set.seed(123)
split <- sample.split(1:dim(Boston)[1], SplitRatio = 0.75)
train.data <- Boston[split,]
test.data <- Boston[!split,]
str(train.data) # 379
str(test.data) # 127
```

Question 2

```{r}
x <- model.matrix(medv~., train.data)[,-1]
y <- train.data$medv
set.seed(123) 
cv <- cv.glmnet(x, y, alpha = 0)
cv$lambda.min
```
Best lambda for ridge regression is 0.6842583

```{r}
model <- glmnet(x, y, alpha = 0, lambda = cv$lambda.min)
coef(model)
```

```{r}
x.test <- model.matrix(medv ~., test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
data.frame(
  RMSE = RMSE(predictions, test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```
RMSE is 4.602301 and coefficient of determination is 0.7063957.

Question 3
```{r}
set.seed(123) 
cv <- cv.glmnet(x, y, alpha = 1)
cv$lambda.min
```

Best lambda for LASSO is 0.0147419

```{r}
model <- glmnet(x, y, alpha = 1, lambda = cv$lambda.min)
coef(model)
```

```{r}
x.test <- model.matrix(medv ~., test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
```

```{r}
data.frame(
  RMSE = RMSE(predictions, test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```
RMSE is 4.66836 and coefficient of determination is 0.6959263.

Question 4
```{r}
set.seed(123)
model <- train(
  medv ~., data = train.data, method = "glmnet",
  trControl = trainControl("cv", number = 10),
  tuneLength = 10
)
model$bestTune
```

```{r}
coef(model$finalModel, model$bestTune$lambda)
```

```{r}
x.test <- model.matrix(medv ~., test.data)[,-1]
predictions <- model %>% predict(x.test)
data.frame(
  RMSE = RMSE(predictions, test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```
RMSE is 4.63037 and coefficient of determination is 0.7010882
